---
title: "R Notebook"
author: "Who, me?"
date: "`r Sys.Date()`"
output:
  html_document:
    df_print: paged
---

# Packages

```{r include=F}
library(tidyverse)
```

# Data

Where does this data come from? What do you want to say about it?

```{r}
starwars
```

# Analysis

What's going on here?

What would you want to say about this?

```{r}
starwars %>%
  count(skin_color) %>%
  arrange(desc(n)) %>%
  slice(1:10) %>%
  ggplot(aes(x = skin_color, y = n, fill = skin_color)) +
  geom_col() +
  labs(
    title = "Top 10 most popular Skin Colors",
    subtitle = "In the Starwars Universe",
    x = "\nSkin Color", y = "Number\n",
    caption = "data source: dplyr::starwars")
```

Explore and try to document what is working and what isn't.

> If you're not sure really where to start and writing this `dplyr` or `ggplot2` code is intimidating you can run these code chunks below to get the hang of things. Try to document what is happenning in the code below.

```{r}
starwars %>% 
  filter(species == "Droid")
```

```{r}
starwars %>% 
  select(name, ends_with("color"))
```

```{r}
starwars %>% 
  mutate(name, bmi = mass / ((height / 100)  ^ 2)) %>%
  select(name:mass, bmi)
```

```{r}
starwars %>% 
  arrange(desc(mass))
```

